---
title: 'Case Study 2: DDS'
author: "Aniketh V"
date: "4/13/2020"
output: html_document
---

#Attrition and Income- Frito Lays

##Introduction
Greetings! Today we will be discussing "Attrition and Income" too check what leads to attrition and income levels for talent management. 

#Setting up the environment
```{r}
#Loading in the required libraries
library(stringr)
library(dplyr)
library(plyr)
library(magrittr)
library(ggplot2)
library(tidyr)
library(GGally)
library(ggalt)
library(mice)
library(e1071)
library(class)
library(caret)

#Reading in the data
attritionData = read.csv("/Users/password1234/Documents/Doing_Data_Science/MSDS_6306_Doing-Data-Science/Unit\ 14\ and\ 15\ Case\ Study\ 2/CaseStudy2-data.csv", header = TRUE)
noAttritionData = read.csv("/Users/password1234/Documents/Doing_Data_Science/MSDS_6306_Doing-Data-Science/Unit\ 14\ and\ 15\ Case\ Study\ 2/CaseStudy2CompSet\ No\ Attrition.csv", header = TRUE)
noSalaryData = read.csv("/Users/password1234/Documents/Doing_Data_Science/MSDS_6306_Doing-Data-Science/Unit\ 14\ and\ 15\ Case\ Study\ 2/CaseStudy2CompSet\ No\ Salary.csv", header = TRUE)

#Get a look at the summary for the attrition data
summary(attritionData)
summary(noAttritionData)
```

##Exploratory Data Analysis
As part of the first step we would like to take a rough look at the data at hand. We want to do some intial analysis to see if anything catches our eye. Additionally, we will take a look at the data types for the variables to see what we are working with.
```{r}
#View the data types for each variable we are dealing with
dplyr::glimpse(attritionData)

#Separate the numerical and categorical variables in the dataframe
sapply(attritionData, is.factor)

#Categorical variables only
categoricalVar = attritionData[,sapply(attritionData, is.factor)] 
summary(categoricalVar)

#Numerical variables only
numericalVar = attritionData[,!sapply(attritionData, is.factor)]
summary(numericalVar)

#Lets look at a rough analysis on multiple different columns to see if we can figure something out
attritionData %>% select(Attrition, JobRole, MonthlyIncome) %>% ggpairs(aes(color = Attrition))
attritionData %>% select(MonthlyIncome, Education, TotalWorkingYears, Attrition) %>% ggpairs(mapping = aes(color = Attrition)) + ggtitle("Monthly Income vs Education, Experience, and Attrition")
attritionData %>% select(MonthlyIncome, YearsAtCompany, TotalWorkingYears, YearsInCurrentRole, OverTime) %>% ggpairs(mapping = aes(color = OverTime)) + ggtitle("Monthly Income vs Years in Industry by Overtime")

#Lets look at if Monthyly income and Job Role play a role in determining Attrition
attritionData %>% ggplot(aes(x = JobRole, y = MonthlyIncome, fill = Attrition)) + geom_bar(stat = "identity", width = .5) + labs(title = "Income", subtitle = "along with Attrition") + theme(axis.text.x = element_text(angle = 65, vjust = 0.6))

attritionData %>% ggplot(aes(x = MonthlyIncome, y = YearsAtCompany, color = JobRole)) + geom_point(position = "jitter")
```
At a glance, it looks like a common variable that influences Attrition as well as monthly income is Total Working Years (experience). There is a moderate correlation between Monthly Income and Years at company along with attrition.

```{r}
#ScatterPlot for trends
attritionData %>% ggplot(aes(x = MonthlyIncome, y = Education)) + geom_point(position = "jitter")

attritionData %>% ggplot(aes(x = Age, y = Attrition)) + geom_point(position = "jitter")


#Need to dummy code Attrition
attritionData$AttritionNum = str_replace(attritionData$Attrition, "No", "0")

attritionData$AttritionNum = str_replace(attritionData$AttritionNum, "Yes", "1")

attritionData$AttritionNum = as.integer(attritionData$AttritionNum)

noAttr = noAttritionData %>% select(ID, WorkLifeBalance, JobInvolvement, YearsAtCompany)

attritionData %>% ggplot(aes(x = Attrition)) + geom_bar()

#Lets do KNN
case = attritionData %>% select(ID, WorkLifeBalance, JobInvolvement, YearsAtCompany, Attrition)
att = data_frame(WorkLifeBalance = 2, JobInvolvement = 1, YearsAtCompany = 5)
knn(case[,c(2,3,4)], att, case$Attrition, k = 3, prob = TRUE)
attr_pred = knn(case[,c(2,3,4)], noAttr[,c(2,3,4)], case$Attrition, k = 19, prob = TRUE)
noAttr$Attrition = attr_pred
table(attr_pred, noAttr$Attrition)
noAttr = noAttr[,-5]

#Train Data and Test Set
set.seed(6)
splitPerc = .70
trainIndices = sample(1:dim(case)[1], round(splitPerc * dim(case)[1]))
train = case[trainIndices,]
test = case[-trainIndices,]

#K = 3
classifications = knn(train[,2:4], test[,2:4], train$Attrition, k = 3, prob = TRUE)
table(classifications, test$Attrition)
confusionMatrix(table(classifications, test$Attrition))
#K = 7
classifications = knn(train[,2:4], test[,2:4], train$Attrition, k = 7, prob = TRUE)
table(classifications, test$Attrition)
confusionMatrix(table(classifications, test$Attrition))
#k=13
classifications = knn(train[,2:4], test[,2:4], train$Attrition, k = 13, prob = TRUE)
table(classifications, test$Attrition)
confusionMatrix(table(classifications, test$Attrition))

#Now going to loop through more than one k and find the average
iterations = 100
set.seed(6)
numks = 30
masterAcc = matrix(nrow = iterations, ncol = numks)
kkk = c()
Sens = c()
Spec = c()
  
for(j in 1:iterations)
{
accs = data.frame(accuracy = numeric(30), k = numeric(30))
trainIndices = sample(1:dim(case)[1], round(splitPerc * dim(case)[1]))
train = case[trainIndices,]
test = case[-trainIndices,]
for(i in 1:numks)
{
  classifications = knn(train[, c(2:4)], test[, c(2:4)], train$Attrition, prob = TRUE, k = i)
  table(classifications, test$Attrition)
  CM = confusionMatrix(table(classifications,test$Attrition))
  masterAcc[j,i] = CM$overall[1]
  kkk[i] = CM$overall[1]
  Sens[i] = CM$byClass[1]
  Spec[i] = CM$byClass[2]
}
}
MeanAcc = colMeans(masterAcc)
plot(seq(1,numks,1),MeanAcc, type = "l")

combo = data_frame(k = 1:30, Accuracy = kkk ,Sensitivity = Sens, Specificity = Spec, MeanAcc)

```



```{r}
#linear models for Salaries (on attrition data before the split)
reg = lm(MonthlyIncome~Education, attritionData)
reg2 = lm(MonthlyIncome~TotalWorkingYears, attritionData)
mReg = lm(MonthlyIncome~TotalWorkingYears + Education, attritionData)

#Test with just two observations
df = data.frame(TotalWorkingYears = 10, Education = 4)
#Create new Data Frame with just experience and education to predict salary for each person
EduTotal = noSalaryData %>% select(ID,TotalWorkingYears, Education)
#Prediction
pred = predict(mReg, newdata = df, interval = "confidence")
pred2 = predict(mReg, newdata = EduTotal)

#New Column with Preds
EduTotal$MonthlySalary = pred2

EduTotal$MonthlySalary = as.integer(EduTotal$MonthlySalary)

#Validation Set
vReg = lm(MonthlySalary~TotalWorkingYears + Education, EduTotal)
vReg

confint(mReg)


#Residual Sum of Squares
RSS = c(crossprod(vReg$residuals))
#Mean Squared Error
MSE = RSS / length(vReg$residuals)
#Root MSE
RMSE = sqrt(MSE)
```

```{r}
#Split data to train and test
set.seed(6)
splitPerc = .70
trainIndices = sample(1:dim(attritionData)[1], round(splitPerc * dim(attritionData)[1]))
train = attritionData[trainIndices,]
test = attritionData[-trainIndices,]

#Linear Reg on Training Model
fit = lm(MonthlyIncome~TotalWorkingYears + Education, train)
#Linear Reg on Testing Model
fit2 = lm(MonthlyIncome~TotalWorkingYears + Education, test)

#Train Residual Sum of Squares
RSS_T = c(crossprod(fit$residuals))
#Train Mean Squared Error
MSE_T = RSS_T / length(fit$residuals)
#Train Root MSE
RMSE_T = sqrt(MSE_T)

#Test Residual Sum of Squares
RSS_Te = c(crossprod(fit2$residuals))
#Test Mean Squared Error
MSE_Te = RSS_Te / length(fit2$residuals)
#Test Root MSE
RMSE_Te = sqrt(MSE_Te)
```